---
title: "Introduction to ggplot2"
subtitle: "(and a little tidyverse)"
author: "Jacob Spangler"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Overview

What is ggplot2?
 - **ggplot2** is a system for declaratively creating graphics, based on The Grammar of Graphics. You provide the data, tell **ggplot2** how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details.

---

# How to install?

```{r eval=F, tidy=F}

# The easiest way to get ggplot2 is to install the whole tidyverse:
install.packages("tidyverse")

# Alternatively, install just ggplot2:
install.packages("ggplot2")

# Or the the development version from GitHub:
# install.packages("devtools")
devtools::install_github("tidyverse/ggplot2")

```

---

# Iris Data

```{r, echo=F}

library(tidyverse)
Iris_Data <- iris %>% as.tbl()

Iris_Data

```

---

# Scatterplot

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length, y=Sepal.Width)) +
  geom_point()

```

---

# Options for Scatterplots

- Add a diagonal line

.pull-left[

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length, y=Sepal.Width)) +
  geom_point() +
  geom_abline()

```

]

.pull-right[

- Add color

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length, y=Sepal.Width,
                      color=Species)) +
  geom_point() +
  geom_abline()

```

]

---

# Histograms

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length)) +
  geom_histogram()

```

---

# Options for Histograms

.pull-left[

- Outline the bins
- Set bin sizes

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length)) +
  geom_histogram(color="black", binwidth = 0.2)

```

]


.pull-right[

- Facet histograms by other variables
- Remove Legend

```{r, fig.height=4, fig.width=4}

ggplot(Iris_Data, aes(x=Sepal.Length, fill=Species)) +
  geom_histogram(color="black", binwidth = 0.2) +
  facet_wrap(~Species, ncol=1) +
  theme(legend.position = "none")

```

]

---

# Dataset

- What about example data set?

```{r, echo=F}

Example_Data <- read_csv("../data/raw/Public Potable Water Systems FINAL 06-22-2018_0.csv",
                         col_types = cols())

Example_Data

```

---

# Scatterplot

```{r, echo=F}

ggplot(Example_Data, aes(x=`Residential Population`, 
                         y=`Number of COMBINED Service Connections (CB)`)) +
  geom_point(alpha=0.4) +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma)

```

---

# Histogram

```{r}

ggplot(Example_Data, aes(x=`Residential Population`)) +
  geom_histogram(color="black") +
  scale_x_continuous(labels = scales::comma)

```

